Gut-resident microorganisms and their genes are associated with cognition and neuroanatomy in children

### 2023-10-11 manuscript/main.tex
pathobionts

We did not identify any statistical associations between individual species
and cognitive performance in our youngest subset (0 to 6 months).
We hypothesize that this may be because brain developmental trajectories
begin \textit{in utero}, and microbial influences (that can only begin at birth)
take additional time to manifest.
In line with this hypothesis, RF models that were able to learn associations
between early-life microbiomes and future cognitive function scores (Figure 3).
The top-ranked species for these models included several 
taxa that were important for predicting concurrent cognitive assessment scores after 18 months, notably
pathobiontic \textit{Enterobacteriaceae} such as
\textit{Escherichia} and \textit{Klebsiella}. 
Representatives of these genera are known to be early and long-term persistent
colonizers of infant intestine that can, in some settings,
promote inflammatory response \cite{popeEnterobacEarlyGut2019}.
By contrast, several inflammation-reducing representatives
of the \textit{Bifidobacterium} genus,
from the early HMO-metabolizer \textit{B. longum} to the
later complex carbohydrate relying \textit{B. pseudocatenulatum} \cite{chungBifidoPsudocatenulatum2021},
commonly associated with weaning to solid food in early age,
also showed high relative importance.
The involvement of most important taxa in different aspects of inflammatory response
resonate with previous findings that associate chronic inflammation with
changes to neural responses in some types of stimulae \cite{xieChronicInfNeuro2019}.

>>>>

We did not identify any statistical associations between individual species
and cognitive performance in our youngest subset (0 to 6 months).
We hypothesize that this may be because brain developmental trajectories
begin \textit{in utero}, and microbial influences (that can only begin at birth)
take additional time to manifest.
In line with this hypothesis, RF models that were able to learn associations
between early-life microbiomes and future cognitive function scores (Figure 3).
The top-ranked species for these models included several 
taxa that were important for predicting concurrent cognitive assessment scores after 18 months, notably
pathobionts \textit{Enterobacteriaceae} such as
\textit{Escherichia} and \textit{Klebsiella}. 
Representatives of these genera are known to be early and long-term persistent
colonizers of infant intestine that can, in some settings,
promote inflammatory response \cite{popeEnterobacEarlyGut2019}.
By contrast, several inflammation-reducing representatives
of the \textit{Bifidobacterium} genus,
from the early HMO-metabolizer \textit{B. longum} to the
later complex carbohydrate relying \textit{B. pseudocatenulatum} \cite{chungBifidoPsudocatenulatum2021},
commonly associated with weaning to solid food in early age,
also showed high relative importance.
The involvement of most important taxa in different aspects of inflammatory response
resonate with previous findings that associate chronic inflammation with
changes to neural responses in some types of stimulae \cite{xieChronicInfNeuro2019}.
